Pulmonary acini exhibit complex changes during postnatal rat lung development

### 2020-12-11 manuscript.tex
Nicer display of DOI of preprint [ci skip]
This build is based on
https://github.com/habi/acinar-analysis-manuscript/commit/a04b78817dad0c1dd6ce05f2e961363c2da567ba.

This commit was created by the following CI build and job:
https://github.com/habi/acinar-analysis-manuscript/commit/a04b78817dad0c1dd6ce05f2e961363c2da567ba/checks
https://github.com/habi/acinar-analysis-manuscript/runs/415877981

A preprint of this manuscript is available on \emph{bioRxiv} at \url{https://doi.org/10.1101/786681} and has been submitted to the \href{https://journals.physiology.org/journal/ajplung}{\emph{American Journal of Physiology - Lung Cellular and Molecular Physiology}}.

This version of the manuscript \href{https://github.com/habi/acinar-analysis-manuscript/compare/d6fc17bd66bdca5cbff30d53e1dbc02d5351609b...9d5c623b3eefb444db71bb9adfe13d2531bc9516}{has been updated} since the first round of review.

This manuscript
(\href{https://habi.github.io/acinar-analysis-manuscript/v/9d5c623b3eefb444db71bb9adfe13d2531bc9516/}{permalink})
was automatically generated
from \href{https://github.com/habi/acinar-analysis-manuscript/tree/9d5c623b3eefb444db71bb9adfe13d2531bc9516}{habi/acinar-analysis-manuscript@9d5c623}
on December 8, 2020.

@@

\begin{itemize}
\item
  \textbf{David Haberthür}
  · \includegraphics{images/orcid.svg}
  \href{https://orcid.org/0000-0003-3388-9187}{0000-0003-3388-9187}
  · \includegraphics{images/github.svg}
  \href{https://github.com/habi}{habi}
  · \includegraphics{images/twitter.svg}
  \href{https://twitter.com/habi}{habi}
  Institute of Anatomy, University of Bern, Switzerland
\item
  \textbf{Eveline Yao}
  Institute of Anatomy, University of Bern, Switzerland
\item
  \textbf{Sébastien F. Barré}
  Institute of Anatomy, University of Bern, Switzerland
\item
  \textbf{Tiziana P. Cremona}
  · \includegraphics{images/orcid.svg}
  \href{https://orcid.org/0000-0003-0061-9578}{0000-0003-0061-9578}
  Institute of Anatomy, University of Bern, Switzerland
\item
  \textbf{Stefan A. Tschanz}
  · \includegraphics{images/orcid.svg}
  \href{https://orcid.org/0000-0003-3880-4437}{0000-0003-3880-4437}
  Institute of Anatomy, University of Bern, Switzerland
\item
  \textbf{Johannes C. Schittny}✉
  · \includegraphics{images/orcid.svg}
  \href{https://orcid.org/0000-0003-4025-3961}{0000-0003-4025-3961}
  Institute of Anatomy, University of Bern, Switzerland
\end{itemize}

@@

To our best knowledge, the present manuscript is the first to quantitatively describe the number of alveoli in individual acini and changes in the volumes of single rat lung acini.
Based on stereological analysis of a large number of individual acini we were able to efficiently extract qualitative and quantitative data of large amounts of single acini from non-destructive three-dimensional tomographic data sets in a highly reproducible manner.

We can easily estimate the total number of acini during lung development from our data.
The limitations of this study are mainly twofold: a) due to technical reasons (i.e.~available beamtime and limitations of the setup we could not image the entire lung and b) lung location (i.e.~ lower medial tip of the right lobe).
However, our sample is more than suitable for the drawn conclusions because it was previously shown that the right lower lobe is a valid sample for the entire lung~(\protect\hyperlink{ref-14OP85b2F}{2}) and that information drawn from one lobe is representative of the whole lung~(\protect\hyperlink{ref-I9TmP6IU}{52}).
Furthermore, based on stereological data, Zeltner et al.~(\protect\hyperlink{ref-I9TmP6IU}{52}) were able to show that the lung parenchyma is quite homogeneous.
Looking at individual acini, Kizhakke Puliyakote et al.~(\protect\hyperlink{ref-RGBeCf8v}{20}) observed that the peripheral, pleura facing acini possess an acinar volume which is two thirds larger than the volume of the central ones.
However, they did not report additional regional differences, e.g.~between acini located cranially and caudally.
Own preliminary data point to the same result.
We imaged an entire rat lung \emph{in situ} immediately post mortem using high-resolution synchrotron radiation based X-ray tomographic microscopy~(\protect\hyperlink{ref-Zv1XcgAO}{4}).
A first analysis of regional difference shows similar results as mentioned above.

@@

By dividing the parenchymal volumes of each lung estimated by Tschanz et al.
(Table 1,~(\protect\hyperlink{ref-wnl86DEM}{46})) by the mean number of acini estimated by Barré et al.~(\protect\hyperlink{ref-uFNlWogb}{3}) we can estimate the mean acinar volume for each day.
The acinus volume we estimated by point counting (Cavalieri estimation) are on average 2.07 times lower but show a similar increase over the studied period (see figure \ref{fig:06}).
It appears that the comparison between data obtained by stereology based on classical paraffin sections or on any kind of 3D-imaging has its limitation if it comes to the absolute numbers.
First, due to the \emph{coast of Britain}-effect the exact value of every surface measurement is depending on the resolution of the imaging method~(\protect\hyperlink{ref-AdXhaEV4}{24}).
Second, the thickness of the alveolar septa in 3D-imaging data is heavily depending on the segmentation between airspace and tissue---regardless if it is done manually or by a computed algorithm.
We observed that the airspace volumes obtained by 3D-imaging has a tendency to be smaller than the ones determined using classical paraffin sections.
However, the relative numbers obtained in one study match the ones of other studies quite well.

\begin{figure}
\hypertarget{fig:06}{%
\centering
\includegraphics{images/fig06.png}
\caption{Plot of the mean acinar volumes.
By dividing the parenchymal volumes of each lung estimated by Tschanz et al.
(Table 1,~(\protect\hyperlink{ref-wnl86DEM}{46})) by the mean number of acini estimated by Barre et al.~(\protect\hyperlink{ref-uFNlWogb}{3}) we can estimate the mean acinar volume for each measured time point (blue plot).
We assessed the volumes of the acini by point counting (Cavalieri estimation, orange plot).
Left: original data.
Right: Our data, scaled by a linear correction factor.
Our volumes of the acini are on average 2.07 times lower but show a similar increase over the studied period.
While the estimation based on Tschanz et al.
is from the entire right middle lobe and on Barre et al.
globally for the whole organ, our estimation is solely based on the stereological analysis of the acini in the lower medial tip of the right lower lung lobe.}\label{fig:06}
}
\end{figure}

@@

Previously, it was shown that the number of acini (e.g.~counted acinar entraces) remains constant during postnatal lung development from day 4 to 60~(\protect\hyperlink{ref-14OP85b2F}{2}), with a mean of 5840~±~547~acini per lung (in another strain of animals, full lung data available for days 4 and 60).
Even though we did not aim to assess the number of acini in this manuscript, calculating this number helps to validate our assessment and helps us to collate the data that can be derived from our stereological analysis to prior results.

@@

In this study, we assessed the number of alveoli per acinus. The alveolar By multiplication of the number of acini by the number of alveoli we estimate a total amount of alveoli per lung of \SIlist{0.64 +- 0.34; 1.23 +- 0.39; 2.65 +-  0.57; 3.01 +- 0.58}{million\  alveoli} for days 4, 10, 21, \& 60, respectively.
Tschanz et al.~estimated these total numbers of alveoli per lung as \SIlist{0.823 +- 0.01; 3.538 +- 1.26; 14.303 +- 3.08; 19.297 +- 3.06}{million\ alveoli} for days 4, 10, 21, \& 60, respectively (\protect\hyperlink{ref-wnl86DEM}{46}), Table 1).
On average, our results are five times smaller but show the same trend when linearly scaled with a factor of 5.04.
The values from Tschanz et al.
overlap well with our linearly scaled values, as shown in figure \ref{fig:08}.

@@

However, relative numbers as reported in this study are more reliable.
This is especially the case when the comparison is done in groups of animals processed in parallel and counted by the same operator.
We assessed the same lungs as Tschanz et al.~(\protect\hyperlink{ref-wnl86DEM}{46}), but based on a very different imaging method and with two different persons performing the stereological assessment.
This may explain the relative factor of 5.04 that we found as a difference.

@@

By dividing the alveolar air space by the number of alveoli (both from Table 1 of~(\protect\hyperlink{ref-wnl86DEM}{46})) we can estimate a mean alveolar volume of 5.91~×~10\textsuperscript{5}~μm\textsuperscript{3} for day 4, of 2.49~×~10\textsuperscript{5}~μm\textsuperscript{3} for day 10, of 1.34~×~10\textsuperscript{5}~μm\textsuperscript{3} for day 21, and 2.93~×~10\textsuperscript{5}~μm\textsuperscript{3} for day 60.
On average, our values (calculated by dividing the mean acinar volume by the mean count of alveoli per acinus) are 1.81 times larger than these estimates of individual alveolar volume (our values day 4: 4.98~×~10\textsuperscript{5}~μm\textsuperscript{3}, day 10: 3.34~×~10\textsuperscript{5}~μm\textsuperscript{3}, day 21 3.07~×~10\textsuperscript{5}~μm\textsuperscript{3} and day 60 8.12~×~10\textsuperscript{5}~μm\textsuperscript{3}), but scale equally during postnatal lung development.
Tschanz et al.
estimated the total alveolar volume without the alveolar ducts.
Since we directly estimated the volume fraction of the alveoli per acinus our number is actually expected to be larger.

@@

To our best knowledge, no in-depth characterization of the acinar properties as presented here---i.e.~spanning the lung development period---was available up to now.
The data presented in this study can contribute to improve computational simulations of pulmonary ventilation and particle deposition.

>>>>

A preprint of this manuscript is available on \emph{bioRxiv} with the DOI \href{https://doi.org/10.1101/786681}{10.1101/786681} and is under revision in the \href{https://journals.physiology.org/journal/ajplung}{\emph{American Journal of Physiology - Lung Cellular and Molecular Physiology}}.
The manuscript \href{https://github.com/habi/acinar-analysis-manuscript/compare/d6fc17bd66bdca5cbff30d53e1dbc02d5351609b...a04b78817dad0c1dd6ce05f2e961363c2da567ba}{has been updated} since the first round of review.

This manuscript
(\href{https://habi.github.io/acinar-analysis-manuscript/v/a04b78817dad0c1dd6ce05f2e961363c2da567ba/}{permalink})
was automatically generated
from \href{https://github.com/habi/acinar-analysis-manuscript/tree/a04b78817dad0c1dd6ce05f2e961363c2da567ba}{habi/acinar-analysis-manuscript@a04b788}
on December 11, 2020.

@@

\begin{itemize}
\item
  \textbf{David Haberthür}
  · \includegraphics{images/orcid.svg}
  \href{https://orcid.org/0000-0003-3388-9187}{0000-0003-3388-9187}
  · \href{mailto:haberthuer@ana.unibe.ch?subject=Feedback\%20regarding\%20the\%20acinus\%20manuscript\%20(a04b788)}{haberthuer@ana.unibe.ch}
  · \includegraphics{images/github.svg}
  \href{https://github.com/habi}{habi}
  · \includegraphics{images/twitter.svg}
  \href{https://twitter.com/habi}{habi}
  Institute of Anatomy, University of Bern, Switzerland
\item
  \textbf{Eveline Yao}
  Institute of Anatomy, University of Bern, Switzerland
\item
  \textbf{Sébastien F. Barré}
  Institute of Anatomy, University of Bern, Switzerland
\item
  \textbf{Tiziana P. Cremona}
  · \includegraphics{images/orcid.svg}
  \href{https://orcid.org/0000-0003-0061-9578}{0000-0003-0061-9578}
  Institute of Anatomy, University of Bern, Switzerland
\item
  \textbf{Stefan A. Tschanz}
  · \includegraphics{images/orcid.svg}
  \href{https://orcid.org/0000-0003-3880-4437}{0000-0003-3880-4437}
  Institute of Anatomy, University of Bern, Switzerland
\item
  \textbf{Johannes C. Schittny}✉
  · \includegraphics{images/orcid.svg}
  \href{https://orcid.org/0000-0003-4025-3961}{0000-0003-4025-3961}
  · \href{mailto:schittny@ana.unibe.ch?subject=Feedback\%20regarding\%20the\%20acinus\%20manuscript\%20(a04b788)}{schittny@ana.unibe.ch}
  Institute of Anatomy, University of Bern, Switzerland
\end{itemize}

@@

To our best knowledge, the present manuscript is the first to quantitatively describe the number of alveoli in individual acini and changes in the volumes of single rat lung acini.
Based on stereological analysis of a large number of individual acini we were able to efficiently extract qualitative and quantitative data from non-destructive three-dimensional tomographic data sets of individual acini in a highly reproducible manner.

We can easily estimate the total number of acini during lung development from our data.
The limitations of this study are mainly twofold: a) due to technical reasons (i.e.~available beamtime and limitations of the setup we could not image the entire lung and b) lung location (i.e.~lower medial tip of the right lobe).
However, our sample is more than suitable for the drawn conclusions because it was previously shown that the right lower lobe is a valid sample for the entire lung~(\protect\hyperlink{ref-14OP85b2F}{2}) and that information drawn from one lobe is representative of the whole lung~(\protect\hyperlink{ref-I9TmP6IU}{52}).
Furthermore, based on stereological data, Zeltner et al.~(\protect\hyperlink{ref-I9TmP6IU}{52}) were able to show that the lung parenchyma is quite homogeneous.
Looking at individual acini, Kizhakke Puliyakote et al.~(\protect\hyperlink{ref-RGBeCf8v}{20}) observed that the peripheral, pleura facing acini possess an acinar volume which is two thirds larger than the volume of the central ones.
However, they did not report additional regional differences, e.g.~between acini located cranially and caudally.
Own preliminary data point to the same result.
We imaged an entire rat lung \emph{in situ} immediately post mortem using high-resolution synchrotron radiation based X-ray tomographic microscopy~(\protect\hyperlink{ref-Zv1XcgAO}{4}).
A first analysis of regional difference shows similar results as mentioned above.

@@

By dividing the parenchymal volumes of each lung estimated by Tschanz et al.~(Table 1,~(\protect\hyperlink{ref-wnl86DEM}{46})) by the mean number of acini estimated by Barré et al.~(\protect\hyperlink{ref-uFNlWogb}{3}) we can estimate the mean acinar volume for each day.
The acinus volume we estimated by point counting (Cavalieri estimation) are on average 2.07 times lower but show a similar increase over the studied period (see figure \ref{fig:06}).
It appears that the comparison between data obtained by stereology based on classical paraffin sections or on any kind of 3D-imaging has its limitation if it comes to the absolute numbers.
First, due to the \emph{coast of Britain}-effect the exact value of every surface measurement is depending on the resolution of the imaging method~(\protect\hyperlink{ref-AdXhaEV4}{24}).
Second, the thickness of the alveolar septa in 3D-imaging data is heavily depending on the segmentation between airspace and tissue---regardless if it is done manually or by a computed algorithm.
We observed that the airspace volumes obtained by 3D-imaging has a tendency to be smaller than the ones determined using classical paraffin sections.
However, the relative numbers obtained in one study match the ones of other studies quite well.

\begin{figure}
\hypertarget{fig:06}{%
\centering
\includegraphics{images/fig06.png}
\caption{Plot of the mean acinar volumes.
By dividing the parenchymal volumes of each lung estimated by Tschanz et al.~(Table 1,~(\protect\hyperlink{ref-wnl86DEM}{46})) by the mean number of acini estimated by Barre et al.~(\protect\hyperlink{ref-uFNlWogb}{3}) we can estimate the mean acinar volume for each measured time point (blue plot).
We assessed the volumes of the acini by point counting (Cavalieri estimation, orange plot).
Left: original data.
Right: Our data, scaled by a linear correction factor.
Our volumes of the acini are on average 2.07 times lower but show a similar increase over the studied period.
While the estimation based on Tschanz et al.~is from the entire right middle lobe and on Barre et al.~globally for the whole organ, our estimation is solely based on the stereological analysis of the acini in the lower medial tip of the right lower lung lobe.}\label{fig:06}
}
\end{figure}

@@

Previously, it was shown that the number of acini (e.g.~counted acinar entraces) remains constant during postnatal lung development from day 4 to 60~(\protect\hyperlink{ref-14OP85b2F}{2}), with a mean of 5840~±~547~acini per lung (in another strain of animals, the data from the whole lungs is available for days 4 and 60).
Even though we did not aim to assess the number of acini in this manuscript, calculating this number helps to validate our assessment and helps us to collate the data that can be derived from our stereological analysis to prior results.

@@

In this study, we assessed the number of alveoli per acinus. The alveolar By multiplication of the number of acini by the number of alveoli we estimate a total amount of alveoli per lung of 0.64 +- 0.34, 1.23 +- 0.39, 2.65 +- 0.57 and 3.01 +- 0.58 million alveoli for days 4, 10, 21, \& 60, respectively.
Tschanz et al.~estimated these total numbers of alveoli per lung as 0.823 +- 0.01, 3.538 +- 1.26, 14.303 +- 3.08 and 19.297 +- 3.06 million alveoli for days 4, 10, 21, \& 60, respectively (\protect\hyperlink{ref-wnl86DEM}{46}), Table 1).
On average, our results are five times smaller but show the same trend when linearly scaled with a factor of 5.04.
The values from Tschanz et al.~overlap well with our linearly scaled values, as shown in figure \ref{fig:08}.

@@

However, relative numbers as reported in this study are more reliable.
This is especially the case when the comparison is done in groups of animals processed in parallel and counted by the same operator.
We assessed the same lungs as Tschanz et al.~(\protect\hyperlink{ref-wnl86DEM}{46}), but performed our assessement based on a very different imaging method and with two different persons performing the stereological assessment at two very different points in time.
This may explain the relative factor of 5.04 that we found as a difference.

@@

By dividing the alveolar air space by the number of alveoli (both from Table 1 of~(\protect\hyperlink{ref-wnl86DEM}{46})) we can estimate a mean alveolar volume of 5.91~×~10\textsuperscript{5}~μm\textsuperscript{3} for day 4, of 2.49~×~10\textsuperscript{5}~μm\textsuperscript{3} for day 10, of 1.34~×~10\textsuperscript{5}~μm\textsuperscript{3} for day 21, and 2.93~×~10\textsuperscript{5}~μm\textsuperscript{3} for day 60.
On average, our values (calculated by dividing the mean acinar volume by the mean count of alveoli per acinus) are 1.81 times larger than these estimates of individual alveolar volume (our values day 4: 4.98~×~10\textsuperscript{5}~μm\textsuperscript{3}, day 10: 3.34~×~10\textsuperscript{5}~μm\textsuperscript{3}, day 21 3.07~×~10\textsuperscript{5}~μm\textsuperscript{3} and day 60 8.12~×~10\textsuperscript{5}~μm\textsuperscript{3}), but scale equally during postnatal lung development.
Tschanz et al.~estimated the total alveolar volume without the alveolar ducts.
Since we directly estimated the volume fraction of the alveoli per acinus our number is actually expected to be larger.

@@

To our best knowledge, no in-depth characterization of the acinar properties as presented here---i.e.~spanning the lung development period---was available up to now.
Even though our results are not precisely accordant with data calculated from previous studies, they match well, found differences are discussed extensively.

The data presented in this study can contribute to improve computational simulations of pulmonary ventilation and particle deposition.
